---
title: "DPLYR penguins"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(here)
```

```{r}
#penguins <- read_csv(here("data/penguins_lter.csv"))
```

```{r}
penguins <- palmerpenguins::penguins
```

## 3. Filter()

#### **Practice Activity 1**

Use `filter()` to create a subset from `penguins` that only contains gentoo penguins with a bill depth greater than or equal to 15.5 millimeters.

```{r}
gentoo_penguins_long_bill <- penguins %>% 
  filter(species == "Gentoo", bill_depth_mm >= 15.5)
```

```{r}
ggplot(penguins, aes(x = bill_depth_mm, fill = species)) +
  geom_histogram() +
  geom_vline(xintercept = 15.5)
```

#### **Practice Activity 2**

Use `filter()` to create a subset from `penguins` that contains observations for male penguins recorded at Dream and Biscoe Islands.

```{r}
penguins %>% 
  filter(sex == "male", island == "Dream" | island == "Bisoce")
```

#### **Practice Activity 3**

Use `filter()` to create a subset from `penguins` that contains observations for female Adelie penguins with bill lengths less than 35 mm.

```{r}
penguins %>% 
  filter(species == "Adelie", sex == "female", bill_length_mm < 35)
```

#### **Practice Activity 4**

Create a subset from `penguins` containing observations for female chinstrap penguins on Dream and Torgersen Islands.

```{r}
penguins %>% 
  filter(sex == "female", species == "Chinstrap", island == "Dream" | island == "Torgersen") 
```

#### **Practice Activity 5**

Create a subset from `penguins` that contains penguins that are either gentoos **OR** have a body mass greater than 4500 g.

```{r}
large_gentoos <- penguins %>% 
  filter(species == "Gentoo" | body_mass_g > 4500)
```

```{r}
min(large_gentoos$body_mass_g, na.rm = TRUE)
```
