---
title: "EDA Assignment Prep"
author: "Amber Camp (edited by Christian Raquepo)"
format: html 
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}
pokedata <- read_csv(here("data/Pokemon.csv"))

pokedata <- pokedata %>% 
  clean_names()
```

```{r}

```

### Describe the data:

Data author: **ALBERTO BARRADAS**

[*https://www.kaggle.com/datasets/abcsds/pokemon?resource=download*](https://www.kaggle.com/datasets/abcsds/pokemon?resource=download)

Description: ***This is pokemon data for 721 pokemon and their stats.***

How many observations (rows) are there? ***800***

How many variables (columns) are there? ***13***

### skimr: Quick overview

```{r}
skim_data <- skimr::skim(pokedata)
```

### Summary: Quick overview

```{r}
summary(pokedata)
```

### Categorical vs. Continuous

```{r}
pokedata$number <- as.factor(pokedata$number)
class(pokedata$number)
summary(pokedata$number)

summary(pokedata$type_1)
pokedata$type_1 <- as.factor(pokedata$type_1)
class(pokedata$type_1)
summary(pokedata$type_1)
```

How many categorical variables are there? ***5/6***

How many continuous variables are there? ***7***

What are some variables of interest to you? ***s***

------------------------------------------------------------------------

## EDA Assignment Prep: **Students Exam Scores**

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}
studentdata <- read.csv(here("data/Scores.csv"))

studentdata <- studentdata %>% 
  clean_names()
```

### Describe the data:

Data author: ***Mr. Royce Kimmons***

Description: ***This is Student Exam Scores data for 1000 students.***

How many observations (rows) are there? 1000.

How many variables (columns) are there? 16.

### Summary: Quick overview

```{r}
skim_data <- skimr::skim(studentdata)
```

### Categorical vs. Continuous

```{r}
summary(pokedata)
```

```{r}
studentdata$x <- as.factor(studentdata$x)
class(studentdata$x)
summary(studentdata$x)

summary(studentdata$ethnic_group)
studentdata$ethnic_group <- as.factor(studentdata$ethnic_group)
class(studentdata$ethnic_group)
summary(studentdata$ethnic_group)
```

How many categorical variables are there? 9

How many continuous variables are there? 6

What are some variables of interest to you? Lunch type, study hours, test prep
